---
title: "Final Presentation"
author: "Chris Kardish and Marie Agosta"
date: "November 18, 2016"
output: pdf_document
---

This is where we clean and prepare the data

```{r}

#rm(list = ls()) this command clears your environment
#setwd('C:/Users/Chris/Documents/GitHub/pair3')
#setwd('C:/Users/Chris/Documents/R/col_soc_science')

library(foreign) 
library(survey)

# R will crash if a primary sampling unit (psu) has a single observation
# set R to produce conservative standard errors instead of crashing
options( survey.lonely.psu = "adjust" )

#import data 
anes <- read.dta("anes_timeseries_2012_Stata12.dta")

#table(anes$dem_age_r_x) #age
anes <- anes[!(anes$dem_age_r_x <= -2 | anes$dem_age_r_x >= 88),] 
#deletes outliers

#table(anes$postvote_rvote)
levels(anes$postvote_rvote) <- c("Refused", "Don't know", "Incomplete", "Nonresponsive", "Missing", "Inapplicable", "No", "No", "No", "Yes")
anes <- anes[(anes$postvote_rvote == "Yes" | anes$postvote_rvote == "No"),] 
anes$postvote_rvote <-  droplevels(anes$postvote_rvote)
anes$postvote_rvote <- as.numeric(anes$postvote_rvote) 
anes$postvote_rvote[anes$postvote_rvote == 1] <- 0
anes$postvote_rvote[anes$postvote_rvote == 2] <- 1
#keeps only Yes/no, changes the variable to a numeric dummy, with 1 as "Yes"

#table(anes$dem_edugroup_x) #education
levels(anes$dem_edugroup_x) <- c("Refused", "Don't know", "Data missing", "Below high school", "High school", "Some post-high", "Bachelor", "Graduate") 
anes <- anes[(anes$dem_edugroup_x == "Below high school" | 
                anes$dem_edugroup_x == "High school" | 
                anes$dem_edugroup_x == "Some post-high" | 
                anes$dem_edugroup_x == "Bachelor" |
                anes$dem_edugroup_x == "Graduate"),] #keeps only these groups
anes$dem_edugroup_x <-  droplevels(anes$dem_edugroup_x)
anes$dem_edugroup_x <- as.numeric(anes$dem_edugroup_x)
anes$dem_edugroup_x <- as.factor(anes$dem_edugroup_x) #gives us 1 to 5, with graduate education
#being the highest

#table(anes$dem_raceeth_x) #race
levels(anes$dem_raceeth_x) <- c("Data missing", "White", "Black", "Asian" , "Hawaiian or Pacif Islr", "Native American or Alaska Native", "Hispanic","Other") 
anes <- anes[!(anes$dem_raceeth_x == "Data missing"),] 
anes$dem_raceeth_x <-  droplevels(anes$dem_raceeth_x)

#table(anes$interest_voted2008) #voting in past election (2008)
levels(anes$interest_voted2008) <- c("Refused", "Don't know", "Yes", "No")
anes <- anes[(anes$interest_voted2008 == "Yes" | anes$interest_voted2008 == "No"),] 
anes$interest_voted2008 <-  droplevels(anes$interest_voted2008)
anes$interest_voted2008 <- as.numeric(anes$interest_voted2008) 
anes$interest_voted2008[anes$interest_voted2008 == 2] <- 0
#turns it into a numeric dummy variable, with 1 meaning "Yes, I voted in 2008."

#table(anes$gender_respondent_x) #gender
anes$female <- anes$gender_respondent_x == "2. Female" #creates dummy gender #variable
anes$female <- as.numeric(anes$female) 
#creates numeric dummy variable, with 1 as "female"

#table(anes$dem_emptype_unemp) #employment status
levels(anes$dem_emptype_unemp) <- c("No", "Yes")
anes$unemployed <- anes$dem_emptype_unemp == "Yes"
anes$unemployed <- as.numeric(anes$unemployed)

#combining needed variables in smaller data set 
anes_small <- data.frame(anes$caseid, anes$dem_age_r_x, anes$dem_edugroup_x, anes$dem_raceeth_x, anes$unemployed, anes$female, anes$interest_voted2008, anes$postvote_rvote, anes$weight_full, anes$psu_full, anes$strata_full)

#		"caseid" ,				      unique record identifiers
		
#		"weight.full" , 		    full-sample weights
		
#		"strata.full" ,			    full-sample strata variable

#		"psu.full" ,			      full-sample cluster variable
		
#		"dem_age_r_x" ,         respondent age group
		
# 	"female" ,	            respondent sex
		
#		"dem_edugroup_x" ,		  how many hours do you work the average week
		
#		"dem_raceeth_x"				  do you support or oppose the tea party   

#   "unemployed"            respondent employment status 

#  interest_voted2008       repsondents previous voting history

#    postvote_rvote          

colnames(anes_small) <- c("caseID", "age", "education", "race", "unemployed", "female", "vote_2008", "vote_2012", "weights", "psu", "strata" )

#set up an object that allows for proper id and weights to be used in the data set

 
#other variable ideas to include, all of which have a lot of observations  
  #table(anes$respons_elections) #efficacy of elections
  #table(anes$preswin_dutychoice_x) #voting a duty 
  #table(anes$pid_x) #strength of party ID
  #table(anes$trust_social) #social trust
  #table(anes$inc_incgroup_pre) #income groups
  #http://stackoverflow.com/questions/9604001/grouping-2-levels-of-a-factor-in-r
  #table(anes$postvote_presstr) #strength of voting preference 

```

Creating the age-band subsets from the smaller ANES data set that contains only 
variables of interest.

```{r}

anes_genY <- subset(anes_small, anes$dem_age_r_x > 17 & anes$dem_age_r_x < 33) 
#creates Millennial subset 
anes_genX <- subset(anes_small, anes$dem_age_r_x > 32 & anes$dem_age_r_x < 48) 
#creates Generation X subset 
anes_boomer <- subset(anes_small, anes$dem_age_r_x > 47 & anes$dem_age_r_x < 67) 
#creates Baby Boomer subset 
anes_silent <- subset(anes_small, anes$dem_age_r_x > 66 & anes$dem_age_r_x < 88) 
#creates Silent Generation subset

ANESdesign_genY <- svydesign(~psu ,  strata = ~strata , data = anes_genY , weights = ~weights, variables = NULL, nest = TRUE)

ANESdesign_genX <- svydesign(~psu ,  strata = ~strata , data = anes_genX , weights = ~weights, variables = NULL, nest = TRUE)

ANESdesign_boomer <- svydesign(~psu ,  strata = ~strata , data = anes_boomer, weights = ~weights, variables = NULL, nest = TRUE)

ANESdesign_silent <- svydesign(~psu ,  strata = ~strata , data = anes_silent, weights = ~weights, variables = NULL, nest = TRUE)
```

Running logistic regression model with confidence intervals

```{r}

#Millennials
M_genY <- svyglm(vote_2012 ~ education + unemployed + female + vote_2008 + race, design = ANESdesign_genY, family = "quasibinomial") 
summary(M_genY)
genY_coef <- coef(M_genY)
exp(genY_coef)

#Baby Boomers
M_boomer <- svyglm(vote_2012 ~ education + unemployed + female + vote_2008 + race, design = ANESdesign_boomer, family = "quasibinomial") 
summary(M_boomer)
boomer_coef <- coef(M_boomer)
exp(boomer_coef)

#Gen X
M_genX <- svyglm(vote_2012 ~ education + unemployed + female + vote_2008 + race, design = ANESdesign_genX, family = "quasibinomial") 
summary(M_genX)
genX_coef <- coef(M_genX)
exp(genX_coef)


# Looking at confidence intervals of parameter point estimates 

confint(M_genY)
confint(M_boomer)
confint(M_genX)

#Giving odds ratios with confidence intervals in a table 

exp(cbind(OR = coef(M_genY), confint(M_genY)))
```

Now playing with and plotting predicted probabilities based on the model

```{r}

#creating a data frame of fitted values 

blackfem_genY <- with(anes_genY, data.frame(female = 1, race = "Black", vote_2008 = 0, unemployed = 0, education = factor(1:5))) 

#predict probability point estimates for each fitted value.

blackfem_genY$predicted_prob <- predict(M_genY, newdata = blackfem_genY, type = "response", se = TRUE)

#the way the code is laid out in the manuel - perhaps we can try this..? I just don't understand most of it 
 predict(object,newdata=NULL,total=NULL,
                             type=c("link","response","terms"),
                             se.fit=(type != "terms"),vcov=FALSE,...)

#so this would be a prediction based on a black, employed Millennial female at
#various levels of education

blackfem_boomer <- with(anes_boomer,
  data.frame(female = 1, race = "Black", vote_2008 = 0, unemployed = 0, education = factor(1:5))) 
blackfem_boomer$prob <- predict(M_boomer, newdata = blackfem_boomer, type = "response")
#so this would be a prediction based on a black, employed Boomer female at
#various levels of education. One thing to notice is that a black female Baby Boomer
#is far less likely to vote at lower levels of education than a Millennial black female,
#based on this model. 

#Now to plot this with confidence intervals.
#First, add the confidence intervals

blackfem_genY <- cbind(blackfem_genY, predict(M_genY, newdata = blackfem_genY, type="link", se=TRUE))
blackfem_genY <- within(blackfem_genY, {
  prob <- plogis(fit)
  LL <- plogis(fit - (1.96 * se.fit))
  UL <- plogis(fit + (1.96 * se.fit))
})

# want to have lines for the two generation we are comparing 
library(ggplot2)
ggplot(blackfem_genY, aes(x = education, y = prob, group = 1)) +
  geom_ribbon(aes(ymin = LL, ymax = UL), alpha = .2) +
  geom_line(aes(colour = education), size=1)

#this plot doesn't look great at the moment, but it does work. 
```

Assessing the model fit. I draw on the UCLA page here for this: 
http://www.ats.ucla.edu/stat/r/dae/logit.htm 

In short: the results below are good, but I don't really know how to explain them.

```{r}

with(M_genY, null.deviance - deviance) #gives chi-square value
with(M_genY, df.null - df.residual) #gives degrees of freed
with(M_genY, pchisq(null.deviance - deviance, df.null - df.residual, lower.tail = FALSE)) 
#last command gives p-value, which is very small and tells us "our model as a whole fits #significantly better than an empty model," according to the UCLA page that Christopher
#used for some of the demonstration in class. But I have no idea what that actually means.
````


Notes from class Nov. 18

Should have correlation matrix like on slide 22 of the Nov. 18 class

Run the simulated stuff he did for presentation purposes. Need a regression table
and other stuff, but for highlighting the main effects that's simply not
a good idea. Simluation helps show the results and how confident we are
about those results. Gives an expectation of what we would see in the "real world"
based on the assumptions of our model. 
