---
title: "Final Presentation"
author: "Chris Kardish and Marie Agosta"
date: "November 18, 2016"
output: pdf_document
---

This is where we clean and prepare the data

```{r}

#rm(list = ls()) this command clears your environment
#setwd('C:/Users/Chris/Documents/GitHub/final_project')
#setwd('C:/Users/Chris/Documents/R/col_soc_science')

# calling appropriate packages 

library(foreign) 
library(survey)

# R will crash if a primary sampling unit (psu) has a single observation
# so we set R to produce conservative standard errors instead of crashing
options( survey.lonely.psu = "adjust" )

#import data from the ANES
anes <- read.dta("anes_timeseries_2012_Stata12.dta")

# to delete ages that do not fit into any of our generations, we delete the unnecesary ages
#table(anes$dem_age_r_x) #age
anes <- anes[!(anes$dem_age_r_x <= -2 | anes$dem_age_r_x >= 88),] 

#We renamed the levels pertaining to if someone voted and then we coded them to be binary, keeping only Yes/no, changes the variable to a numeric dummy, with 1 as "Yes"
#table(anes$postvote_rvote)
levels(anes$postvote_rvote) <- c("Refused", "Don't know", "Incomplete", "Nonresponsive", "Missing", "Inapplicable", "No", "No", "No", "Yes")
anes <- anes[(anes$postvote_rvote == "Yes" | anes$postvote_rvote == "No"),] 
anes$postvote_rvote <-  droplevels(anes$postvote_rvote)
anes$postvote_rvote <- as.numeric(anes$postvote_rvote) 
anes$postvote_rvote[anes$postvote_rvote == 1] <- 0
anes$postvote_rvote[anes$postvote_rvote == 2] <- 1

#We renamed the levels pertaining to education and recoded them as numeric ranging from 
#1-5 (1 being below high school) and 5 (graduate education)
#table(anes$dem_edugroup_x) #education
levels(anes$dem_edugroup_x) <- c("Refused", "Don't know", "Data missing", "Below high school", "High school", "Some post-high", "Bachelor", "Graduate") 
anes <- anes[(anes$dem_edugroup_x == "Below high school" | 
                anes$dem_edugroup_x == "High school" | 
                anes$dem_edugroup_x == "Some post-high" | 
                anes$dem_edugroup_x == "Bachelor" |
                anes$dem_edugroup_x == "Graduate"),] #keeps only these groups
anes$dem_edugroup_x <-  droplevels(anes$dem_edugroup_x)
anes$dem_edugroup_x <- as.numeric(anes$dem_edugroup_x)

#We renamed the labels pertaining to race and removed all missing data 
#table(anes$dem_raceeth_x) #race
levels(anes$dem_raceeth_x) <- c("Data missing", "White", "Black", "Asian" , "Hawaiian or Pacif Islr", "Native American or Alaska Native", "Hispanic","Other") 
anes <- anes[!(anes$dem_raceeth_x == "Data missing"),] 
anes$dem_raceeth_x <-  droplevels(anes$dem_raceeth_x)
anes$dem_raceeth_x <- as.numeric(anes$dem_raceeth_x)

#white = 1, black = 2, asian = 3, hawaiin/pacific islander = 4, native american = 5
#hispanic = 6

#We renamed the labels pertaining to voting in the 2008 election, dropped NA ones, and recoded #the remaining yes and no to be 1 or 0. 
#table(anes$interest_voted2008) #voting in past election (2008)
levels(anes$interest_voted2008) <- c("Refused", "Don't know", "Yes", "No")
anes <- anes[(anes$interest_voted2008 == "Yes" | anes$interest_voted2008 == "No"),] 
anes$interest_voted2008 <-  droplevels(anes$interest_voted2008)
anes$interest_voted2008 <- as.numeric(anes$interest_voted2008) 
anes$interest_voted2008[anes$interest_voted2008 == 2] <- 0

#We created a dummy variable for gender, with 1 as "female"
#table(anes$gender_respondent_x) #gender
anes$female <- anes$gender_respondent_x == "2. Female" #creates dummy gender #variable
anes$female <- as.numeric(anes$female) 

#We renamed the levels pertaining to employment status 
#table(anes$dem_emptype_unemp) #employment status
levels(anes$dem_emptype_unemp) <- c("No", "Yes")
anes$unemployed <- anes$dem_emptype_unemp == "Yes"
anes$unemployed <- as.numeric(anes$unemployed)

#To simplify our task we create a smaller data set that includes all relevant variables
#, which are listed below the code. 

anes_small <- data.frame(anes$caseid, anes$dem_age_r_x, anes$dem_edugroup_x, anes$dem_raceeth_x, anes$unemployed, anes$female, anes$interest_voted2008, anes$postvote_rvote, anes$weight_full, anes$psu_full, anes$strata_full)

#		"caseid" ,				      unique record identifiers
		
#		"weight.full" , 		    full-sample weights
		
#		"strata.full" ,			    full-sample strata variable

#		"psu.full" ,			      full-sample cluster variable
		
#		"dem_age_r_x" ,         respondent age group
		
# 	"female" ,	            respondent sex
		
#		"dem_edugroup_x" ,		  how many hours do you work the average week
		
#		"dem_raceeth_x"				  do you support or oppose the tea party   

#   "unemployed"            respondent employment status 

#  interest_voted2008       repsondents previous voting history

#    postvote_rvote         ???


# We named the variables so they are very easy to understand 
colnames(anes_small) <- c("caseID", "age", "education", "race", "unemployed", "female", "vote_2008", "vote_2012", "weights", "psu", "strata" )


#other variable ideas to include, all of which have a lot of observations  
  #table(anes$respons_elections) #efficacy of elections
  #table(anes$preswin_dutychoice_x) #voting a duty 
  #table(anes$pid_x) #strength of party ID
  #table(anes$trust_social) #social trust
  #table(anes$inc_incgroup_pre) #income groups
  #http://stackoverflow.com/questions/9604001/grouping-2-levels-of-a-factor-in-r
  #table(anes$postvote_presstr) #strength of voting preference 

```

Creating the age-band subsets from the smaller ANES data set that contains only 
variables of interest.

```{r}

anes_genY <- subset(anes_small, anes$dem_age_r_x > 17 & anes$dem_age_r_x < 33) 
#creates Millennial subset 
anes_genX <- subset(anes_small, anes$dem_age_r_x > 32 & anes$dem_age_r_x < 48) 
#creates Generation X subset 
anes_boomer <- subset(anes_small, anes$dem_age_r_x > 47 & anes$dem_age_r_x < 67) 
#creates Baby Boomer subset 
anes_silent <- subset(anes_small, anes$dem_age_r_x > 66 & anes$dem_age_r_x < 88) 
#creates Silent Generation subset

#set up an objects to run with the logit models that allow for results to be 
#statistically representative of the population through weighting 

ANESdesign_genY <- svydesign(~psu ,  strata = ~strata , data = anes_genY , weights = ~weights, variables = NULL, nest = TRUE)

ANESdesign_genX <- svydesign(~psu ,  strata = ~strata , data = anes_genX , weights = ~weights, variables = NULL, nest = TRUE)

ANESdesign_boomer <- svydesign(~psu ,  strata = ~strata , data = anes_boomer, weights = ~weights, variables = NULL, nest = TRUE)

ANESdesign_silent <- svydesign(~psu ,  strata = ~strata , data = anes_silent, weights = ~weights, variables = NULL, nest = TRUE)
```

Running logistic regression model with confidence intervals

```{r}

#Running a logit, generation by generation. Then looking at a summary of the results, follows by looking at and creating an object for the coeffs of the results. Then converting those results to probabilities by exp(ing) them 

#Millennials
M_genY <- svyglm(vote_2012 ~ education + unemployed + female + vote_2008 + race, design = ANESdesign_genY, family = "quasibinomial") 
summary(M_genY)
genY_coef <- coef(M_genY)
exp(genY_coef)

#Baby Boomers
M_boomer <- svyglm(vote_2012 ~ education + unemployed + female + vote_2008 + race, design = ANESdesign_boomer, family = "quasibinomial") 
summary(M_boomer)
boomer_coef <- coef(M_boomer)
exp(boomer_coef)

#Gen X
M_genX <- svyglm(vote_2012 ~ education + unemployed + female + vote_2008 + race, design = ANESdesign_genX, family = "quasibinomial") 
summary(M_genX)
genX_coef <- coef(M_genX)
exp(genX_coef)


#Looking at confidence intervals of parameter point estimates 

confint(M_genY)
confint(M_boomer)
confint(M_genX)

#Giving odds ratios with confidence intervals in a table 

exp(cbind(OR = coef(M_genY), confint(M_genY)))
```

Now playing with and plotting predicted probabilities based on the model

```{r}

#Manually collecting probabilities

####Black Millennial Woman
#1-probability for below-high-school educated black employed woman who didn't vote in 2008
prob_genY_ed1 <- (exp(-0.3448 - 0.3448 + 0.1249 + 0.6389))/(1+exp(-0.3448 - 0.3448 + 0.1249 + 0.6389)) 
#2-probability for high-school educated black employed woman who didn't vote in 2008
prob_genY_ed2 <- (exp(-0.3448 + 0.5263 + 0.1249 + 0.6389))/(1+exp(-0.3448 + 0.5263 + 0.1249 + 0.6389)) 
#3-probability for some post-high-school educated black employed woman who didn't vote in 2008
prob_genY_ed3 <- (exp(-0.3448 + 0.7955 + 0.1249 + 0.6389))/(1+exp(-0.3448 + 0.7955 + 0.1249 + 0.6389)) 
#4-probability for bachelor-educated black employed woman who didn't vote in 2008
prob_genY_ed4 <- (exp(-0.3448 + 0.8990 + 0.1249 + 0.6389))/(1+exp(-0.3448 + 0.8990 + 0.1249 + 0.6389)) 
#5-probability for graduate-educated black employed woman who didn't vote in 2008
prob_genY_ed5 <- (exp(-0.3448 + 1.5099 + 0.1249 + 0.6389))/(1+exp(-0.3448 + 1.5099 + 0.1249 + 0.6389)) 


####Black Baby Boomer Woman
#1-probability for below-high-school educated black employed woman who didn't vote in 2008
prob_boomer_ed1 <- (exp(-0.87481 - 0.87481 + 0.22553 -0.76681))/(1+exp(-0.87481 - 0.87481 + 0.22553 -0.7668)) 
#2-probability for high-school educated black employed woman who didn't vote in 2008
prob_boomer_ed2 <- (exp(-0.87481 + 0.81340 + 0.22553 -0.76681))/(1+exp(-0.87481 + 0.81340 + 0.22553 -0.76681)) 
#3-probability for some post-high-school educated black employed woman who didn't vote in 2008
prob_boomer_ed3 <- (exp(-0.87481 + 1.12412 + 0.22553 -0.76681))/(1+exp(-0.87481 + 1.12412 + 0.22553 -0.76681)) 
#4-probability for bachelor-educated black employed woman who didn't vote in 2008
prob_boomer_ed4 <- (exp(-0.87481 + 1.69306 + 0.22553 -0.76681))/(1+exp(-0.87481 + 1.69306 + 0.22553 -0.76681)) 
#5-probability for graduate-educated black employed woman who didn't vote in 2008
prob_boomer_ed5 <- (exp(-0.87481 + 1.54687 + 0.22553 -0.76681))/(1+exp(-0.87481 + 1.54687 + 0.22553 -0.76681)) 


predicted_fit <- function(alpha, beta1, beta2, beta3, beta4, beta5, 
    x1, x2, x3, x4, x5, 
    se1, se2, se3, se4, se5)
{
lb_fun <- function(coef, se){
    lb <- coef - (1.96 * se)
    return(lb)
}

ub_fun <- function(coef, se){
    ub <- coef + (1.96 * se)
    return(ub)
}

fit <- alpha + beta1 * x1 + beta2 * x2 + beta3 * x3 + beta4 * x4 + beta5 * x5
lb <- alpha + lb_fun(beta1, se1) * x1  + lb_fun(beta2, se2) * x2  + lb_fun(beta3, se3) * x3
ub <- alpha + ub_fun(beta1, se1) * x1  + ub_fun(beta2, se2) * x2  + ub_fun(beta3, se3) * x3
predictions <- c(lb, fit, ub)

probs <- function(x) {(exp(x)) / (1 + exp(x)) }
predictions <- probs(predictions)
return(predictions) 
}

predicted_fit(-0.17336, 0.23787, -0.43932, 0.08985, 0.96896, -0.01822, 1, 0, 1, 0, 2, 0.11959, 0.37259, 0.21844, 0.22869, 0.06800)



#creating a dataframe of predicted variables for plotting and other uses
blackfem_genY_ed <- c(prob_genY_ed1, prob_genY_ed2, prob_genY_ed3, prob_genY_ed4, prob_genY_ed5)
blackfem_boomer_ed <- c(prob_boomer_ed1, prob_boomer_ed2, prob_boomer_ed3, prob_boomer_ed4, prob_boomer_ed5)
education <- cbind(blackfem_genY_ed, blackfem_boomer_ed)
rownames(education) <- c("Below high school", "High school", "Some post-high", "Bachelor", "Graduate")
colnames(education) <- c("Millennial probability", "Baby Boomer probability")

#creating a data frame of fitted values for a hypothetical black woman
blackfem_genY <- with(anes_genY, data.frame(female = 1, race = '2', vote_2008 = 0, unemployed = 0, education = factor(1:5))) 

#this is an alternative we've played around with that includes the weighting columns
#used by the survey package as well
blackfem_genY <- with(anes_genY, data.frame(caseID = factor(1:5), weights = .506, psu = 1, strata = 6, female = 1, race = 2, vote_2008 = 0, unemployed = 0, education = factor(1:5))) 

#predict probability point estimates for each fitted value.

blackfem_genY$predicted_prob <- prediction(M_genY, data = blackfem_genY, type = "response")

blackfem_genY$predicted_prob <- predict(M_genY, newdata = blackfem_genY, type = "response", se = TRUE)
#another alternative
blackfem_genY$predicted_prob <- predict(M_genY, newdata = blackfem_genY, type = "response", 
                                        design = ANESdesign_genY)

#Now to plot this with confidence intervals.
#First, add the confidence intervals

blackfem_genY <- cbind(blackfem_genY, predict(M_genY, newdata = blackfem_genY, type="link", se=TRUE))
blackfem_genY <- within(blackfem_genY, {
  prob <- plogis(fit)
  LL <- plogis(fit - (1.96 * se.fit))
  UL <- plogis(fit + (1.96 * se.fit))
})

# want to have lines for the two generation we are comparing 
library(ggplot2)
ggplot(blackfem_genY, aes(x = education, y = prob, group = 1)) +
  geom_ribbon(aes(ymin = LL, ymax = UL), alpha = .2) +
  geom_line(aes(colour = education), size=1)

```

Assessing the model fit. I draw on the UCLA page here for this: 
http://www.ats.ucla.edu/stat/r/dae/logit.htm 

In short: the results below are good, but I don't really know how to explain them.

```{r}

with(M_genY, null.deviance - deviance) #gives chi-square value
with(M_genY, df.null - df.residual) #gives degrees of freed
with(M_genY, pchisq(null.deviance - deviance, df.null - df.residual, lower.tail = FALSE)) 
#last command gives p-value, which is very small and tells us "our model as a whole fits #significantly better than an empty model," according to the UCLA page that Christopher
#used for some of the demonstration in class. But I have no idea what that actually means.
````


Notes from class Nov. 18

Should have correlation matrix like on slide 22 of the Nov. 18 class

Run the simulated stuff he did for presentation purposes. Need a regression table
and other stuff, but for highlighting the main effects that's simply not
a good idea. Simluation helps show the results and how confident we are
about those results. Gives an expectation of what we would see in the "real world"
based on the assumptions of our model. 
